---
title: "MyFirstRmd"
author: "Shaun"
date: "11/5/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYPD Shooting Project

This document contains the analysis of NYPD Shooting Data taken from <https://catalog.data.gov/dataset>
The data contains crime incidents that occured in New York from 2016 to end of 2021, with information on Location (Boro, Latitude, Longitude, Location_Desc) as well as the descriptive detail of the incident such as Victim and Suspect Profile.

Analysis will be done on the Boro to understand the distribution of the crime as well as trying to identify trend in crime using the historical data.



# Importing
```{r Importing, echo=TRUE}
library(tidyverse)
library(lubridate)
url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
nypd_shooting <- read_csv(url)

spec(nypd_shooting)
```

# Tidying
```{r tidying, echo=TRUE}
nypd_shooting <- nypd_shooting %>%
  select(INCIDENT_KEY, OCCUR_DATE, BORO, STATISTICAL_MURDER_FLAG, PERP_AGE_GROUP, PERP_SEX, PERP_RACE, VIC_AGE_GROUP, VIC_SEX, VIC_RACE) %>%
  mutate(OCCUR_DATE  = mdy(OCCUR_DATE ))
```




```{r Data, echo=TRUE}
summary(nypd_shooting)



summary(is.na(nypd_shooting))
```


```{r grouping, echo=TRUE}
OCCUR_DATE_summary <- nypd_shooting %>%
  count(OCCUR_DATE)

BORO_summary <- nypd_shooting %>%
  count(BORO)

VIC_RACE_summary <- nypd_shooting %>%
  count(VIC_RACE)

```



# Analysis & Visualisation
```{r boro_summary, echo=TRUE}
OCCUR_DATE_summary

BORO_summary

VIC_RACE_summary
```

```{r group_date_to_month, echo=TRUE}
OCCUR_DATE_summary <- OCCUR_DATE_summary %>%
  group_by(month=floor_date(OCCUR_DATE, "month")) %>%
  summarize(n = sum(n))
```






```{r visualisation, echo=TRUE}
OCCUR_DATE_summary %>%
  ggplot( aes(x = month, y = n)) + 
  geom_line(aes(color = "n")) +
  ggtitle("Monthly Crime Trend") 
```






```{r test, echo=FALSE}
ggplot(BORO_summary, aes(x = BORO, y = n)) + geom_bar(stat="identity")  +
  ggtitle("Crime per Boro") 
```

```{r brooklyn, echo = TRUE}
nypd_shooting$Brooklyn <- ifelse(nypd_shooting$BORO == 'BROOKLYN',1,0)

nypd_shooting
```


```{r model, echo=TRUE}
mod.boro = lm(Brooklyn~PERP_RACE+PERP_AGE_GROUP, data = nypd_shooting)
summary(mod.boro)
```


```{r session, echo=TRUE}
sessionInfo()
```


# Summary
Brooklyn had the highest crime rate, followed by Bronx, then Queens. Crime rate in New York has been decreasing gradually until the significant 3 month spike in June 2020.

There was no statistically significant finding on Boro and Murder flag, which show potential bias on my personal assumption that there will be a high correlation of murder from more dangerous boro.

There was however statistal significantly on Brooklyn in regards to perpetrator's age and race.



